Estimating Real-Time Highstreet Footfall from Wi-Fi Probe Requests

### 2018-11-03 review/response.md

First review comments
=====================

Title : Estimating Real-Time Highstreet Footfall from Wi-Fi Probe Requests
Manuscript id : IJGIS-2018-0340
Comments dated : 28-Oct-2018

1 Editor
--------

1.1 Overall Comments
---------------------

1. While this is an important and timing topic dealing with data uncertainty
and privacy in the mobile age, the reviewers have raised concerns on
	- Clarity of the methodology
	- Lack of discussion of wifi-based human activity studies in the big data age
	- Proper use of terminology
	- Performance of the algorithm etc.

2. In lieu of these comments, they have recommended to make substantial revisions
to your manuscript

3. Also, since this is a "smart cities" related special issue, I ask you to
	- Add more discussions on smart cities in the introduction and 
	  conclusion sections and
	- Link your research more tightly to this theme.

>>> Thank you for the very helpful comments of the reviewers of our paper/
>>> As requested, we have responded to all comments without adding
>>> significant length to the paper. We have also extended our discussion of
>>> how effective measurement of footfall is relevant to understanding the
>>> functioning of the Smart City.
>>> 
>>> If any of what follows and what has been uploaded is unclear, please do
>>> not hesitate to contact me.
>>> 
>>> Yours sincerely
>>> Balamurugan Soundararaj
>>> (on behalf also of Paul Longley and James Cheshire)

1.2 General Instructions
------------------------

1. Please try to avoid making the paper longer (keep in mind that the average
length of papers in this journal is 18-20 published pages). It is important that
you ensure that the spelling and grammar is of a high standard.

2. Please also highlight the changes to your manuscript within the document by
using the track changes mode in MS Word or by using bold or colored text.

3. To revise your manuscript, please select the link below,
https://mc.manuscriptcentral.com/ijgis?URL_MASK=a32caf3bbd49412cac8d1cfa9d7ebaf8
which will take you your decision letter, and where you need to
resubmit your manuscript.

4. If you have used the on-line LaTeX template, upload all those files and a pdf.

5. Please check your English spelling and grammar.

6. There are two hard limits that you need to bear in mind, 
	- Image size - Please be aware that the total number of pixels 
	  (height X width) in each image must be less than 40,000,000 (40 mp).
	- Total file size - please make sure that the combined, uncompressed,
	  file sizes do not exceed 20Mb. The internal LZW compression, which 
	  is an option in TIF, is acceptable however.
	- Please do not save B/W or Greyscale images as RGB.

7. Equations and tables need to be in an editable format.

8. You need to check that the paper title in ScholarOne is the same as the
title on the paper and that the Title page contains all authors’ names and
affiliations as you would want them to appear.

9. When submitting your revised manuscript, you will be able to respond to
the comments made by the reviewers in the space provided.  You can use this
space to document any changes you make to the original manuscript.  In order
to expedite the processing of the revised manuscript, please be as specific as
possible in your response to the reviewers.

10. Please delete any redundant files before completing the submission.

2 Reviewer 1
------------

2.1 Overall Summary
-------------------

This paper is about the estimation of pedestrian footfall using Wi-Fi
signal emitted by mobile devices.
It deals with the problem of estimating the footfall size based on a noisy
and biased data. More specifically it addresses the following issues that may
lead to overcounting:
 - MAC randomization that may cause a device to generate several MAC addresses;
 - Signals received from outside the interest zone; unequal adoption
   of Wi-Fi smartphone in the population.
To address this issue, a number of techniques are introduced:
 - clustering received signal strength to remove device that are outside
   the monitored area;
 - clustering probes based on sequence number and timing;
 - finally an adjustment factor is applied to correct the estimated value.
   The adjustment factor is calibrated for each location using
   a manual counting.
The author conducted a pilot experiment to identify the clustering 
algorithms and parameters, and then they conducted a second range of 
experiments at different location over an extended period of time. During 
each of those experiments, ground-truth count was collected manually.
The results show that the presented approach can significantly reduce the
counting error compared to a naive solution consisting in counting the 
distinct number of MAC observed.
Nevertheless, there are cases (2) where the error is still large (50%),
even if in other cases (1) the error is only of 9%.

>>> This is a correct surmisal of the remit of the paper. We have also added#
>>> material to the introduction and conclusion of the paper to better link
>>> this remit to the Smart Cities discourse.

2.2 General Comments
--------------------

1. The idea of clustering frames using sequence number has been introduced 
in [1]. This work needs to be cited when introducing the proposed clustering
algorithm.
[1] Mathy Vanhoef, Célestin Matte, Mathieu Cunche, Leonardo Cardoso, Frank
Piessens. Why MAC Address Randomization is not Enough: An Analysis of Wi-Fi
Network Discovery Mechanisms. ACM AsiaCCS, May 2016, Xi’an, China. 2016,

>>> Thank you. We have included a citiation to the source, as well as one 
>>> other recent addition to this fast-developing literature.

2. The author considered that the hashing of MAC address is a good way to 
protect users privacy, as some MAC address collected by the device may not be 
random. However hashing of MAC address cannot be considered as a proper 
anonymization technique (see [2] and 
http://webpolicy.org/2014/03/19/questionable-crypto-in-retail-analytics/ ); 
at best it can be considered as an obfuscation. 
[2] Levent Demir, Mathieu Cunche, Cédric Lauradoux. Analysing the privacy 
policies of Wi-Fi trackers. Workshop on Physical Analytics, Jun 2014, Bretton 
Woods, United States. ACM, 2014

>>> Thank you for pointing this out. In the revised version of the paper
>>> we have made reference to this discussion, while also making clear the 
>>> anonymisation is ensured by the system architecture that processess and
>>> curates the data. Spedicifically the weekly rotation of the random salt
>>> with which the hashing process is carried out.

3. There is a number of commercial solutions doing counting based on Wi-Fi 
signal. Although the detail of their solution is not public, it would be 
worthwhile to acknowledge their existence.

>>> Thank you - this has been done.

4. In section 3.2, it is explained that data is divided into two sets 
corresponding on random and non-random MAC. Which method is used to perform 
this classification ?

>>> The identification of a MAC address as randomised or non-randomised is done
>>> looking at the second character of the vendor part of the MAC address.
>>> The addresses with E, A, 2, 6 at these locations are identified to be
>>> randomised MAC addresses. This has been included in the revised paper.

5. How was the sensor positioned compared to the area monitored used for 
ground truth ? A more detailed description is required regarding the positioning
of the sensor and the size and topology of the surveyed area.

>>> This is a good point, and is very important for our validation work. We
>>> have included a more detailed and explicit discussion of the way that field
>>> observation was used to calibrate and cross validate the Wi-Fi results.

2.3 Clarifications & Corrections
--------------------------------

1. The topic addressed in this work is interesting and timely. The proposed
approach is sound and shows promising results. The fact that the authors 
conducted field experiments to evaluate their solution is a positive point. 
Yet before publication, I believe that there are a number of points that needs 
to be clarified or corrected. Regarding the technical aspects I have a
number of minor remarks:

>>> Thank you very much for these very positive comments. The paper reports on
>>> extensive methodological research and field implementation and is, we 
>>> believe, cutting edge in its finding. It also has important implications 
>>> for our understanding of the functioning of smart cities, and these
>>> implications are discussed in greater detail in the revised version.

2. "WiFi beacon technologies to access the Internet". In Wi-Fi, beacons are
management frames used by Access Point for service discovery. They do not
provide Internet access. "Wi-Fi Access Point to access ..."

>>> Thank you. We have corrected and clarified the point.

3. "Wi-Fi antennae regularly broadcast a special type of signal" -> the antenna
is just the part ensuring that the signal is correctly emitted, but it is not 
the source. In this context it is better to use a more general concept and refer
to the "Wi-Fi network interface" (or Wi-Fi interface) as the the element
broadcasting a signal.

>>> Thank you. Point clarified in the revised version.

4. The IEEE 802.11 b/g is just a subpart of the 802.11 specification. Here you 
can refer to the 802.11 specifications as a whole : [“802.11-2012 - IEEE 
Standard for Information technology–Telecommunications and information exchange
between systems Local and metropolitan area networks–Specific requirements 
Part 11: Wireless LAN Medium Access Control (MAC) and Physical Layer (PHY) 
Specifications,” IEEE Std 802.11-2012 (Revision of IEEE Std 802.11-2007),
pp. 1–2793, Mar. 2012.

>>> Thank you. Noted and added to the revised version.

5. "the first step in establishing a Wi-Fi based connection " is also known as
service discovery

>>> Thank you. Added to the revised version

6. "Wi-Fi functionality has been turned off by the user", regarding this point
you may be interested in :

7. "Media Access Control (MAC) address which is an unique identifier for the 
wireless hardware of the mobile device"  The transmitter field of the 802.11
frame can indeed contain a unique MAC address  but can also contain a random 
one as noted by the author. Here it would be better to avoid stating that the 
content of this field is always a unique MAC.

>>> Thank you. Again, this has been noted in the revised version of the paper.

8. "The strength of the signal which transmitted the request.". In fact the 
Received Signal Strength Indicator, a value providing an indication on the 
strength of the received signal as seen by the receiver, not the strength at
the emitter side.

>>> Thank you. This distinction has been noted in the revised version of the
>>> paper.

9. Could you provide more information on the application used for counting 
(Clicker)

>>> This simple application was developed specifically for the project for
>>> Android platform. More details on the application has been included in 
>>> the revised manuscript.

10. "Though the recycling of sequence number r after 4000 leads to multiple
classifications reported on single device, the magnitude of error is greatly
reduced." This sentence is not clear. And the cycle length is not 4000 but 
4096 = 2^12.

>>> Thank you. This point has been boted and accommodated into the revised
>>> version.

11. Regarding the issue of cycling sequence number, one could use a distance 
with a modulo : di,j = di - dj mod 4096.

>>> Thank you very much for the suggestion. From out tests on using this 
>>> method for distance calculation results in a change of --% of the result.
>>> Which in our opinion doesnt influence the results in any significant way
>>> considering the context.

2.4 Typos and other remarks
--------------------------------

1. Challanges

2. "These are deemed to be consumer data because devices carried by consumers 
routinely probe for a consumer service, specifically a Wi-Fi connection.
Monitoring the probes from such devices provides an indication" This sentence 
is not clear.

3. "Though Wi-Fi F is a ‘location-less’ technology, ", please clarify what 
you mean by that.

4. "privacy infringement" may not be the best word as infrigement relate to a 
law or a regulation. Privacy breach or leak may be a better choice.

5. Wireshark and not WireShark.

6. "We hypothesize that an adjustment factor could be arrived at for each 
location of data collection, ""This calibration can be carried over periodically
and the frequency to improve the quality of the estimation." something wrong in 
both sentences.

>>> Thak you for picking up these minor lapses. All have been corrected or 
>>> clarified in the revised version of the paper.

3 Reviewer 2
------------

3.1 Overall Summary
-------------------

1. This paper addressed the data quality issue of Wi-Fi-based probe requests in 
estimating pedestrian counts. The authors proposed several methods to improve
the accuracy of estimating the number of unique mobile devices from a set of 
anonymized probe requests without revealing their original device information.

2. This is a crucial topic in mobile data uncertainty, and the manuscript has 
the potential of becoming a useful reference in the field. However, it was not 
organized effectively and missing many details in the methodology.  Before
addressing these problems, it is difficult to assess the reliability and 
scientific merit of the case study.

3.2 General Comments
--------------------

1. First, the literature review section adequately covered Wi-Fi-based human 
activity studies, but it is in lack of an overview of modeling human activity 
in the big data era. I would suggest the authors add an overarching paragraph 
(or a sub-section) in Section 2 discussing the importance and challenges of 
modeling human activities and urban dynamics based on various types of big 
geodata, such as mobile phone records, location-based social media, Bluetooth 
data, etc.

2. The conclusion did not address most of the important results from the case study and should be expanded.

3.3 Corrections and Clarifications
----------------------------------

Second, the methodology should be better clarified. There are many details 
missing or in need of a better justification. For example,

1. P4 l50, without providing any explanation, the authors hypothesized that 
sequence number and length of the packet are sufficient to estimate the number 
of unique devices. This argument was not elaborated until 2.5 pages later 
in section 3.2.

2. P6 l27, please provide more information regarding how you defined the 
threshold for low/high signal strength when eliminating the background noise. 
This was also unclear in the case study on P8, where the authors repeatedly 
mixed up classification and clustering. Neither k-means nor hierarchical 
clustering is considered a “classification algorithm” as claimed in p8 l26. 
Very few details were provided regarding the parameter setting of 
these analyses.

3. P6 l30-31, in the phone shop example, the authors mentioned that it is 
possible to identify background noise based on a sharp rise of the request 
number; however, this will also eliminate the regular pedestrian flow in that 
shop and cause additional inaccuracy. Please clarify.

4. P7 l21, how often does the recycling of sequence number occur, and how will 
this impact the reliability of your analysis? Please provide more details.

5. P7 Section 3.3, please be more specific about the calibration process and 
the external source of information to be used here.

6. P7 l45, the analysis in this research is based on one single sensor for each 
location; however, in reality, it is very common to estimate travel flows based 
on multiple sensors (e.g., nearby stores in a shopping mall). How would your 
method address duplication in multiple sensors?

7. P10 l48, how was the adjustment factor determined?

4 Reviewer 3
------------

4.1 Overall Summary
-------------------

1. This paper presents Wi-Fi based footfall counting methodology, which is timely and interesting topic. The method can be performed to estimate human activity such as pedestrian footfall from Wi-Fi probe requests.

4.2 General Comments
--------------------

Some areas where I would like to see more detail:

1. It would have been interesting to understand the temporal characteristics for
a week pedestrian footfall. How to consider this situation in your study? Why do
you choose time range from 12:30 to 13:00 hrs ?

2. Will you give more detail information about the data and the parameters for 
various clustering algorithms? Including the number of records, the number of 
clusters.

4.3 Corrections and Clarifications
----------------------------------

More specific comments can be found in the following.

1. Minor formatting issues and typos needs to be fixed.

2. Figure 2, no value and units in the x-axis and y-axis.

3. Table 1, and Figure 4 are in wrong location.

4. Related work: How about WiFi-based real-time data analytics?

5. Methodology: Will you draw a diagram of the Methodology?

4. Performance Analysis: How about the performance for collecting data using 
PostgreSQL? Did you compare Apache Flink and Spark Streaming for real-time data 
collecting?

7. How about the performance for different clustering methods in your study？

8. The reference list needs to be written in a standard format. Some reference papers are lacking specific pages .Please use a standard way of reference list.

### 2018-11-05 review/response.md

1. While this is an important and timing topic dealing with data uncertainty
and privacy in the mobile age, the reviewers have raised concerns on
	- Clarity of the methodology
	- Lack of discussion of wifi-based human activity studies in the big data age
	- Proper use of terminology
	- Performance of the algorithm etc.

2. In lieu of these comments, they have recommended to make substantial revisions
to your manuscript

3. Also, since this is a "smart cities" related special issue, I ask you to
	- Add more discussions on smart cities in the introduction and 
	  conclusion sections and
	- Link your research more tightly to this theme.

> Thank you for the very helpful comments of the reviewers of our paper/
> As requested, we have responded to all comments without adding
> significant length to the paper. We have also extended our discussion of
> how effective measurement of footfall is relevant to understanding the
> functioning of the Smart City.
> 
> If any of what follows and what has been uploaded is unclear, please do
> not hesitate to contact me.
> 
> Yours sincerely
> Balamurugan Soundararaj
> (on behalf also of Paul Longley and James Cheshire)

@@

2 Reviewer 1
------------

2.1 Overall Summary
-------------------

This paper is about the estimation of pedestrian footfall using Wi-Fi
signal emitted by mobile devices.
It deals with the problem of estimating the footfall size based on a noisy
and biased data. More specifically it addresses the following issues that may
lead to overcounting:
 - MAC randomization that may cause a device to generate several MAC addresses;
 - Signals received from outside the interest zone; unequal adoption
   of Wi-Fi smartphone in the population.
To address this issue, a number of techniques are introduced:
 - clustering received signal strength to remove device that are outside
   the monitored area;
 - clustering probes based on sequence number and timing;
 - finally an adjustment factor is applied to correct the estimated value.
   The adjustment factor is calibrated for each location using
   a manual counting.
The author conducted a pilot experiment to identify the clustering 
algorithms and parameters, and then they conducted a second range of 
experiments at different location over an extended period of time. During 
each of those experiments, ground-truth count was collected manually.
The results show that the presented approach can significantly reduce the
counting error compared to a naive solution consisting in counting the 
distinct number of MAC observed.
Nevertheless, there are cases (2) where the error is still large (50%),
even if in other cases (1) the error is only of 9%.

> This is a correct surmisal of the remit of the paper. We have also added#
> material to the introduction and conclusion of the paper to better link
> this remit to the Smart Cities discourse.

@@

2.2 General Comments
--------------------

1. The idea of clustering frames using sequence number has been introduced 
in [1]. This work needs to be cited when introducing the proposed clustering
algorithm.
[1] Mathy Vanhoef, Célestin Matte, Mathieu Cunche, Leonardo Cardoso, Frank
Piessens. Why MAC Address Randomization is not Enough: An Analysis of Wi-Fi
Network Discovery Mechanisms. ACM AsiaCCS, May 2016, Xi’an, China. 2016,

> Thank you. We have included a citiation to the source, as well as one 
> other recent addition to this fast-developing literature.

@@

2. The author considered that the hashing of MAC address is a good way to 
protect users privacy, as some MAC address collected by the device may not be 
random. However hashing of MAC address cannot be considered as a proper 
anonymization technique (see [2] and 
http://webpolicy.org/2014/03/19/questionable-crypto-in-retail-analytics/ ); 
at best it can be considered as an obfuscation. 
[2] Levent Demir, Mathieu Cunche, Cédric Lauradoux. Analysing the privacy 
policies of Wi-Fi trackers. Workshop on Physical Analytics, Jun 2014, Bretton 
Woods, United States. ACM, 2014

> Thank you for pointing this out. In the revised version of the paper
> we have made reference to this discussion, while also making clear the 
> anonymisation is ensured by the system architecture that processess and
> curates the data. Spedicifically the weekly rotation of the random salt
> with which the hashing process is carried out.

@@

3. There is a number of commercial solutions doing counting based on Wi-Fi 
signal. Although the detail of their solution is not public, it would be 
worthwhile to acknowledge their existence.

> Thank you - this has been done.

4. In section 3.2, it is explained that data is divided into two sets 
corresponding on random and non-random MAC. Which method is used to perform 
this classification ?

> The identification of a MAC address as randomised or non-randomised is done
> looking at the second character of the vendor part of the MAC address.
> The addresses with E, A, 2, 6 at these locations are identified to be
> randomised MAC addresses. This has been included in the revised paper.

@@

5. How was the sensor positioned compared to the area monitored used for 
ground truth ? A more detailed description is required regarding the positioning
of the sensor and the size and topology of the surveyed area.

> This is a good point, and is very important for our validation work. We
> have included a more detailed and explicit discussion of the way that field
> observation was used to calibrate and cross validate the Wi-Fi results.

@@

2.3 Clarifications & Corrections
--------------------------------

1. The topic addressed in this work is interesting and timely. The proposed
approach is sound and shows promising results. The fact that the authors 
conducted field experiments to evaluate their solution is a positive point. 
Yet before publication, I believe that there are a number of points that needs 
to be clarified or corrected. Regarding the technical aspects I have a
number of minor remarks:

> Thank you very much for these very positive comments. The paper reports on
> extensive methodological research and field implementation and is, we 
> believe, cutting edge in its finding. It also has important implications 
> for our understanding of the functioning of smart cities, and these
> implications are discussed in greater detail in the revised version.

@@

7. "Media Access Control (MAC) address which is an unique identifier for the 
wireless hardware of the mobile device"  The transmitter field of the 802.11
frame can indeed contain a unique MAC address  but can also contain a random 
one as noted by the author. Here it would be better to avoid stating that the 
content of this field is always a unique MAC.

> Thank you. Again, this has been noted in the revised version of the paper.

8. "The strength of the signal which transmitted the request.". In fact the 
Received Signal Strength Indicator, a value providing an indication on the 
strength of the received signal as seen by the receiver, not the strength at
the emitter side.

> Thank you. This distinction has been noted in the revised version of the
> paper.

@@

9. Could you provide more information on the application used for counting 
(Clicker)

> This simple application was developed specifically for the project for
> Android platform. More details on the application has been included in 
> the revised manuscript.

@@

10. "Though the recycling of sequence number r after 4000 leads to multiple
classifications reported on single device, the magnitude of error is greatly
reduced." This sentence is not clear. And the cycle length is not 4000 but 
4096 = 2^12.

> Thank you. This point has been boted and accommodated into the revised
> version.

@@

11. Regarding the issue of cycling sequence number, one could use a distance 
with a modulo : di,j = di - dj mod 4096.

> Thank you very much for the suggestion. From out tests on using this 
> method for distance calculation results in a change of --% of the result.
> Which in our opinion doesnt influence the results in any significant way
> considering the context.

@@

4. "privacy infringement" may not be the best word as infrigement relate to a 
law or a regulation. Privacy breach or leak may be a better choice.

5. Wireshark and not WireShark.

6. "We hypothesize that an adjustment factor could be arrived at for each 
location of data collection, ""This calibration can be carried over periodically
and the frequency to improve the quality of the estimation." something wrong in 
both sentences.

> Thank you for picking up these minor lapses. All have been corrected or 
> clarified in the revised version of the paper.

@@

3 Reviewer 2
------------

3.1 Overall Summary
-------------------

1. This paper addressed the data quality issue of Wi-Fi-based probe requests in 
estimating pedestrian counts. The authors proposed several methods to improve
the accuracy of estimating the number of unique mobile devices from a set of 
anonymized probe requests without revealing their original device information.

> This is correct, although the intended emphasis upon disclosure is a feature
> rather than a principal motivation. We have also added further discussion of
> how these principally technical issues relate to smart city data
> infrastructure issues.

### 2018-11-05 review/response.md

1. This paper addressed the data quality issue of Wi-Fi-based probe requests in 
estimating pedestrian counts. The authors proposed several methods to improve
the accuracy of estimating the number of unique mobile devices from a set of 
anonymized probe requests without revealing their original device information.

> This is correct, although the intended emphasis upon disclosure is a feature
> rather than a principal motivation. We have also added further discussion of
> how these principally technical issues relate to smart city data
> infrastructure issues.

2. This is a crucial topic in mobile data uncertainty, and the manuscript has 
the potential of becoming a useful reference in the field. However, it was not 
organized effectively and missing many details in the methodology.  Before
addressing these problems, it is difficult to assess the reliability and 
scientific merit of the case study.

> Thank you very much for this endorsement of the motivation for our paper.
> We hope that following responses and the amendments to the paper will allay
> your concerns about the methodology and application.

@@

3.2 General Comments
--------------------

1. First, the literature review section adequately covered Wi-Fi-based human 
activity studies, but it is in lack of an overview of modeling human activity 
in the big data era. I would suggest the authors add an overarching paragraph 
(or a sub-section) in Section 2 discussing the importance and challenges of 
modeling human activities and urban dynamics based on various types of big 
geodata, such as mobile phone records, location-based social media, Bluetooth 
data, etc.

2. The conclusion did not address most of the important results from the case 
study and should be expanded.

> This is a welcome comment, and we have added material in the
> introduction and conclusion as to the value of new Big Data sources for
> developing activity-based representations of behaviour and circumstances
> in the smart city, This chimes with much of our related research, but we
> write cognisant of the article length restrictions of IJGIS.

@@

1. P4 l50, without providing any explanation, the authors hypothesized that 
sequence number and length of the packet are sufficient to estimate the number 
of unique devices. This argument was not elaborated until 2.5 pages later 
in section 3.2.

> This material has now been amended and resequenced.

2. P6 l27, please provide more information regarding how you defined the 
threshold for low/high signal strength when eliminating the background noise. 
This was also unclear in the case study on P8, where the authors repeatedly 
mixed up classification and clustering. Neither k-means nor hierarchical 
clustering is considered a “classification algorithm” as claimed in p8 l26. 
Very few details were provided regarding the parameter setting of 
these analyses.

> Thank you. The reference to clustering has been amended. The
> specifics of the thresholds used have been included, and the techniques
> justified in more detail.

@@

3. P6 l30-31, in the phone shop example, the authors mentioned that it is 
possible to identify background noise based on a sharp rise of the request 
number; however, this will also eliminate the regular pedestrian flow in that 
shop and cause additional inaccuracy. Please clarify.

> Thank you. Clarified in the revised version. It is know with preliminary 
> investigation that the received signal strength indicator decreases with
> increasing distance between the mobile device and the base station. This
> change is also found to be exponential as opposed to linear. With this context
> When we install a sensor to have a clear view of the pedestrian footpath
> in front of it, we make sure that all the devices beyond a certain distance
> will report similarly low signal strength. This threhold is idenfied using
> the kmeans algorithm to find the natural break in the data. We have found that
> this successfully isolates a farther stationary source of large number of
> probe requests such as a phone shop or a restaurant next door, Bustops etc.
> This doesn't work when the source of error is within the range of the
> footfall. As shown in our examples 2,4. Here our methodology fails to provide
> sufficient accuracy.

@@

4. P7 l21, how often does the recycling of sequence number occur, and how will 
this impact the reliability of your analysis? Please provide more details.

> The recycling of the sequence number occurs every 4096 counts. From our
> experiments we observed that a device staying next to device actively searching
> recycles the sequence number 1-2 times in a 30 mins interval. On average
> this is expected to be much less than this. when the device doesn't probe
> as frequently. We have included more details regarding the cycling of sequence
> numbers and have mentioned method to overcomet this as mentioned by other
> reviewer. The change in implementing the method is estimated to be --% which
> we found to be not significant in our context.

@@

5. P7 Section 3.3, please be more specific about the calibration process and 
the external source of information to be used here.

> This section has been expanded to include more details. The calibration process
> is adjusting the future counts based on a sample manual counts done at the
> location. The external source of information used is the manual counts which is
> collected outside the sensor

### 2018-11-05 review/response.md

5. P7 Section 3.3, please be more specific about the calibration process and 
the external source of information to be used here.

> This section has been expanded to include more details. The calibration process
> is adjusting the future counts based on a sample manual counts done at the
> location. The external source of information used is the manual counts which is
> collected outside the sensor

6. P7 l45, the analysis in this research is based on one single sensor for each 
location; however, in reality, it is very common to estimate travel flows based 
on multiple sensors (e.g., nearby stores in a shopping mall). How would your 
method address duplication in multiple sensors?

> This is a valid concern. Which we tried to clarify in the revised manuscript.
> Our motivation is not to attempt to track the movements of individuals,
> although we do attempt to accommodate measurement variations attributable to 
> micro-site location of the sensors. In related work we have carried out 
> controlled comparison of sensor measures obtained from adjacent store
> locations, and examined correspondence with retail unit function. This enables
> identification of the differences in dwell time between, for example,
> fast food outlets and clothes shops. This full analysis requires
> addition of store till data and the analysis is detailed. For this
> reason we include only a short reference to this extension of our
> analysis here.

@@

7. P10 l48, how was the adjustment factor determined?

> The adjustment factor is calculated as the ration between the manual counts to
> the counts calculated from sensor for the sample period. This 
> is used to adjust future counts at the same locations.

### 2018-11-05 review/response.md

5. P7 Section 3.3, please be more specific about the calibration process and 
the external source of information to be used here.

> This section has been expanded to include more details. The calibration process
> is adjusting the future counts based on a sample manual counts done at the
> location. The external source of information used is the manual counts which is
> collected outside the sensor

6. P7 l45, the analysis in this research is based on one single sensor for each 
location; however, in reality, it is very common to estimate travel flows based 
on multiple sensors (e.g., nearby stores in a shopping mall). How would your 
method address duplication in multiple sensors?

> This is a valid concern. Which we tried to clarify in the revised manuscript.
> Our motivation is not to attempt to track the movements of individuals,
> although we do attempt to accommodate measurement variations attributable to 
> micro-site location of the sensors. In related work we have carried out 
> controlled comparison of sensor measures obtained from adjacent store
> locations, and examined correspondence with retail unit function. This enables
> identification of the differences in dwell time between, for example,
> fast food outlets and clothes shops. This full analysis requires
> addition of store till data and the analysis is detailed. For this
> reason we include only a short reference to this extension of our
> analysis here. We also aim this method to count footfall accurately at 
> locations rather than the flow of footfall between those locations. But as 
> demonstarted in
> Lee J, Nemati S, Silva I, Edwards BA, Butler JP, Malhotra A. Transfer entropy 
> estimation and directional coupling change detection in biomedical time series.
> Biomed Eng Online. 2012;11:19. Published 2012 Apr 13.
> doi:10.1186/1475-925X-11-19
> and
> Vicente R, Wibral M, Lindner M, Pipa G. Transfer entropy--a model-free 
> measure of effective connectivity for the neurosciences.
> J Comput Neurosci. 2010;30(1):45-67.
> we can infer the relation or flow between the locations using other
> probabilistic approaches.

### 2018-11-05 review/response.md

1. First, the literature review section adequately covered Wi-Fi-based human 
activity studies, but it is in lack of an overview of modeling human activity 
in the big data era. I would suggest the authors add an overarching paragraph 
(or a sub-section) in Section 2 discussing the importance and challenges of 
modeling human activities and urban dynamics based on various types of big 
geodata, such as mobile phone records, location-based social media, Bluetooth 
data, etc.

> This is a welcome comment, and we have added material in the
> introduction and conclusion as to the value of new Big Data sources for
> developing activity-based representations of behaviour and circumstances
> in the smart city, This chimes with much of our related research, but we
> write cognisant of the article length restrictions of IJGIS.

2. The conclusion did not address most of the important results from the case 
study and should be expanded.

> Yes, we agree, and some wider conclusions about the implications of
> our research for better understanding the functioning of the smart city
> have now been added.

@@

4.2 General Comments
--------------------

Some areas where I would like to see more detail:

1. It would have been interesting to understand the temporal characteristics for
a week pedestrian footfall. How to consider this situation in your study? Why do
you choose time range from 12:30 to 13:00 hrs ?

> Thank you. Our related work examines daily, weekly and seasonal
> variation, and wil be the subject of future papers. There was no reason for
> selecting the time in the pilot study. the location was chosen for it it known
> to have high number of randomised probe requests and has one of the highest
> footfall in London. For the main study the counting was done such that each
> location has atleast 2/3 counts in the same time interval.

@@

2. Will you give more detail information about the data and the parameters for 
various clustering algorithms? Including the number of records, the number of 
clusters.

> Thank you. This has been corrected in the revised version of the
> paper.

@@

3. Table 1, and Figure 4 are in wrong location.

4. Related work: How about WiFi-based real-time data analytics?

5. Methodology: Will you draw a diagram of the Methodology?

> We have resequenced some of the material in response to Referee 2
> and, cognisant of over-all length requirements, think that this (rather
> than a diagram) addresses your concerns.

### 2018-11-05 review/response.md

Some areas where I would like to see more detail:

1. It would have been interesting to understand the temporal characteristics for
a week pedestrian footfall. How to consider this situation in your study? Why do
you choose time range from 12:30 to 13:00 hrs ?

> Thank you. Our related work examines daily, weekly and seasonal
> variation, and wil be the subject of future papers. There was no reason for
> selecting the time in the pilot study. the location was chosen for it it known
> to have high number of randomised probe requests and has one of the highest
> footfall in London. For the main study the counting was done such that each
> location has atleast 2/3 counts in the same time interval.

2. Will you give more detail information about the data and the parameters for 
various clustering algorithms? Including the number of records, the number of 
clusters. How about the performance for different clustering methods 
in your study？

> Thank you. This part has been expanded and clarified in the revised manuscript. The
> performance of the algorithms were tested on the basis of the time taken for
> processing and the relative improvement in results. We found the k-means gives
> us the best data for the least processing time. The graph based clustering
> was done using parameters which were arrived at using trial and error
> controlling for a known device.

@@

4.3 Corrections and Clarifications
----------------------------------

More specific comments can be found in the following.

1. Minor formatting issues and typos needs to be fixed.
Figure 2, no value and units in the x-axis and y-axis.
Table 1, and Figure 4 are in wrong location.

> Thank you. This has been corrected in the revised version of the

2. Related work: How about WiFi-based real-time data analytics?

> We have included reference to such work in our extended discussion of
> the relevance of the work to the smart cities discourse.

3. Methodology: Will you draw a diagram of the Methodology?

@@

> We have resequenced some of the material in response to Referee 2
> and, cognisant of over-all length requirements, think that this (rather
> than a diagram) addresses your concerns.

4. Performance Analysis: How about the performance for collecting data using 
PostgreSQL? Did you compare Apache Flink and Spark Streaming for real-time data 
collecting?

> This is an interesting point in the context of big data and tools for tackling
> big data. Though the data set at hand is large enough for desktop computing to
> be unfeasible, it is neither big nor fast enough for a large scale big data
> systems. A toolkit with opensource, free text processing tools and relational
> database management system was enough to handle the compexity, size and speed
> of the data. The toolkit employed here involves PostgreSQL, Bash, Unix tools
> for text processing (sed, awk, sort, join), NodeJs and R counting.
> This was a proven toolkit used in the larger project (smart street sensor)
> providing near real-time delivery of data.

5. The reference list needs to be written in a standard format. Some reference 
papers are lacking specific pages .Please use a standard way of reference list.

> Thank you. This has been corrected in the revised version of the paper.

### 2018-11-14 paper/sections/abstract.tex
wednesday modifications

The accurate measurement of human activity with high spatial and temporal granularity is crucial for understanding the structure and function of the built environment.
With increasing mobile ownership, the Wi-Fi `probe requests' generated by mobile devices can act as a cheap, scalable and real-time source of data for establishing such measures.
The two major challanges we face in using these probe requests for estimating human activity are, filtering the noise generated by the uncertain field of measurement and clustering anonymised probe requests generated by the same devices together without compromising the privacy of the users.
In this paper we demonstrate that we can overcome these challanges by using class intervals and a novel graph based technique for filtering and clustering the probe requests which in turn, enables us to reliably measure real-time pedestrian footfall at retail highstreets.

>>>>

The accurate measurement of human activity with high spatial and temporal granularity is crucial for understanding the structure and function of the built environment.
With increasing mobile ownership, the Wi-Fi `probe requests' generated by mobile devices can act as a cheap, scalable and real-time source of data for establishing such measures.
The two major challenges we face in using these probe requests for estimating human activity are, filtering the noise generated by the uncertain field of measurement and clustering anonymised probe requests generated by the same devices together without compromising the privacy of the users.
In this paper we demonstrate that we can overcome these challenges by using class intervals and a novel graph based technique for filtering and clustering the probe requests which in turn, enables us to reliably measure real-time pedestrian footfall at retail highstreets.

### 2018-11-14 paper/sections/introduction.tex
wednesday modifications

In this paper, through a set of experiments, we evaluate the value of data collected from a Great Britain wide network of 800 devices \citep{sss2016}, installed in order to characterize the footfall patterns of a scientifically balanced sample of retail centres.
These devices are located in shop windows and record the probes emitted by mobile phones and other Wi-Fi enabled devices.
These are deemed to be consumer data because devices carried by consumers routinely probe for a consumer service, specifically a Wi-Fi connection.
Monitoring the probes from such devices provides an indication of the presence of their users, regardless of whether or not internet connectivity is established.
Our core motivation is to appraise the usefulness of Wi-Fi probe requests harvested from our network of sensors in order to indicate levels of pedestrian activity.
More broadly still, in our future research we intend to classify the nationwide network of footfall profiles as part of a programme of research to understand the form and functioning of retail areas at a time of far-reaching structural change for the retail industry.

>>>>

In this paper, through a set of experiments, we evaluate the value of data collected from a Great Britain wide network of 800 devices \citep{sss2016}, installed in order to characterize the footfall patterns of a scientifically balanced sample of retail centres.
These devices are located in shop windows and record the probes emitted by mobile phones and other Wi-Fi enabled devices.
The data collected from these devices are deemed to be consumer data because devices carried by consumers routinely probe for Wi-Fi connection which is a consumer service.
Monitoring the probes from such devices provides an indication of the presence of their users, regardless of whether or not internet connectivity is established.
Our core motivation is to appraise the usefulness of Wi-Fi probe requests harvested from our network of sensors in order to indicate levels of pedestrian activity.
More broadly still, in our future research we intend to classify the nationwide network of footfall profiles as part of a programme of research to understand the form and functioning of retail areas at a time of far-reaching structural change for the retail industry.
